---
title: "607 Project 2 Bus Routes"
author: "Lu Beyer"
output:
  html_document: default
  pdf_document: default
---

# Overview

This is an extremely large dataset with over 650,000 entries logging bus breakdowns and delays for New York City schools. I want to organize the data in ways that can help me determine the route difficulty of these NYC school buses.  Upon first looking at this data, I am interested in what years have the most breakdowns and delays, which routes have the most breakdowns and delays, and the biggest cause of breakdown or delays on these routes.  I'm considering exploring which boros are most impacted as well, and what the length of delays tend to be, but I will have a clearer idea of how I want to analyze the data once I begin to clean the dataset.


```{r}
library(tidyverse)
```

```{r}
data <- read.csv("Bus_Breakdown_and_Delays_20240225.csv")
```

```{r}
a <- data %>% 
  select(School_Year, Route_Number, Busbreakdown_ID, Reason, Boro, Bus_Company_Name, How_Long_Delayed, Breakdown_or_Running_Late)
```
```{r}
b <- a %>% 
  filter(Route_Number != "") %>% 
  # filter(Reason == "Heavy Traffic") %>%
  filter(School_Year == "2017-2018") %>% 
  group_by(Reason) %>%
  summarise(count = n())
```

```{r}
c <- a %>% 
  select(School_Year, Reason)
```


```{r}
table(c)
```


Because I am working with such a large dataset, I wanted to focus on a single year and look at that data more closely.  I decided to organize my data so I would be able to determine which year had the most delays and breakdowns.  Grouping by year and getting the count of entries per year allowed me to see which year had the most number of break downs and delays documented.  


```{r}
summary <- a %>% 
  group_by(School_Year) %>% 
  summarise(count = n())
```

Once I determined 2018 had the most delays and breakdowns in my dataset, I wanted to see which route number had the most entries.  Filtering by school year and grouping by route number, I was able to get the counts for each route.  Because there were over 6000 bus routes, I wanted to focus on the 10 routes with the highest number of delays and breakdowns so I would be able to create clearer visualizations for that data.
```{r}
Year_2018_2019 <- a %>% 
  filter(School_Year == "2018-2019") %>% 
  group_by(Route_Number) %>%
  summarise(count = n()) %>% 
  arrange(.,count) %>% 
  tail(10)
  
  
```

```{r}
d <- a %>% 
  filter(Route_Number != "") %>% 
  group_by(Reason, School_Year) %>%
  summarise(count = n())
```


```{r}
ggplot(data = Year_2018_2019, aes(x = Route_Number, y = count)) +
  geom_bar(position = "dodge", stat = "identity")
```
I wanted to look at what the biggest cause for delays in 2018 and compare that to the cause of delays over all of the years documented in this dataset.

```{r}
ggplot(data = b, aes(x = count, y = Reason)) +
  geom_bar(position = "dodge", stat = "identity")
```

```{r}
ggplot(data = d, aes(x = count, y = Reason)) +
  geom_bar(position = "dodge", stat = "identity")
```

Finally, I decided to look at which years had the most documented incidents.


```{r}
ggplot(data = summary, aes(x = School_Year, y = count)) +
  geom_bar(position = "dodge", stat = "identity")
```

# Conclusion

There is more that can be done with this data set, but with the variables I chose to analyze, I was able to determine which year had the highest number of incidents (2018), which 10 bus routes within that year were most impacted (1, 2, 3, 5, M326, M817, M922, M965, M978, and M984), and what was the highest cause for delay reported (heavy traffic).  If I wanted to delve deeper into the data I chose to analyze, I could look at the specific route each route number takes to see what parts of the city they run through.  If heavy traffic is the main cause of bus delays as this data shows, it's likely these routes run through areas of the city that are more densely populated or have higher rates of commuters in general.  It could be interesting to see this data to be able confirm this, or see what other causes impacts these traffic on these specific routes.  Additionally, looking at the reported incidents by year, I see that 2020-2021 had by far the lowest reported incidents.  Knowing that this was the year of the start of the pandemic and many schools shifted to remote learning at that time, this dip in incidents makes a lot of sense, and as the 2024 school year has not yet ended, the lower count of reported incidents will likely change. 

I would be interested to look at population data throughout 2015 - 2019 to see if there was an increase in NYC population to cause the steady increase of incidents over that time (given that heavy traffic is the main cause of incidents).  It could also be helpful to look at commuter data to see if there is a trend of more drivers vs public transition commuters.  The dip in incidents between 2019 and 2022 makes sense to me because of the pandemic, but I would be curious to know the reasoning for the trends in the data otherwise. 